---
title: "Week 2 Homework"
format: html
editor: visual
execute: 
  message: false
  warning: false
---

# Rowing Records by Weight

Our goal is to verify if this relationship still holds:

$$V_{max} \propto N^{1/9} $$

## Clean Data

Data is from a wikipedia entry: [List of World Best Times in Rowing](https://en.wikipedia.org/wiki/List_of_world_best_times_in_rowing)

Let's import and clean the data

```{r}
library(tidyverse)
library(lubridate)

week <- "Week_2"
raw_data <- read_csv(here::here(week, "Data",  "rowing_records_2022.csv"))

clean_data <- raw_data |> 
  mutate(Boat = tolower(Boat)) |> 
  mutate(num_rowers = case_when(  # recoding the boat as number of rowers
    str_detect(Boat, "single") ~ 1,
    str_detect(Boat, "pairs|double") ~ 2,
    str_detect(Boat, "four|quad") ~ 4,
    str_detect(Boat, "eight") ~ 8)) |> 
  mutate(time_min = as.numeric(seconds(Time)/3600)) |> 
      mutate(km_h = 2*60/time_min) # let's use the same units 
                                        # as the plot in the assignment
  
  
```
## Visualize Relationships

Let's see what rowers plotted against time looks like on a log-log plot.

```{r}
clean_data |> 
  ggplot(aes(num_rowers, km_h, col = Sex)) + geom_point() + geom_smooth(method = "lm") + 
  scale_x_log10() + scale_y_log10() + theme_light() + 
  labs(title = "World Records for Rowing, Men and Women", subtitle = "Pulled from wikipedia", caption = paste0("As of ", Sys.Date()))

ggsave(here::here(week, "Output", "men_women_fit_rowing.png"))
```
## Model

Let's fit a linear model on the log-log data:

```{r}

models <- clean_data |> 
  group_nest(Sex) |> 
  mutate(coeffs = map(data, ~ lm(log10(km_h) ~ log10(num_rowers), .) |> broom::tidy())) |> 
  ungroup() |> 
  select(Sex, coeffs) |> 
  unnest(coeffs)

```

## Explore Model Results

```{r}

models |>
  select(Sex:estimate) |>
  pivot_wider(names_from = "term", values_from = "estimate") |>
  rename(c = 2, beta = 3) |>
  mutate(c = 10^c) |> ## we need to exponentiate because the coefficient from the model is on the log scale
  knitr::kable()
```

## Interpretation

```{r}
#| output: false

mod_clean <- models |>
  select(Sex:estimate) |>
  pivot_wider(names_from = "term", values_from = "estimate") |>
  rename(c = 2, beta = 3) |>
  mutate(c = 10^c) 

coeff_sex <- mod_clean |> 
  pull(beta, Sex)
```


This gives us beta coeff of `r coeff_sex[["Men"]]` for men and `r coeff_sex[["Women"]]` 
for women. This is fairly close to 1/9 (`r 1/9`).
