---
title: "Week 1 Homework"
format:
   gfm:
     html-math-method: webtex

## General Approach

We are looking to estimate the beta and c coefficients for the exponential model:

$${M}_{WorldRecord} = c{M}_{WeightClass}^{\gamma}$$
Taking the log of both sides, we end up with a linear version of this model:

$$log({M}_{WorldRecord}) = log(c) + {\gamma}log({M}_{WeightClass})$$
We can estimate this using linear regression with the log-log transformation by
taking the log of weight class and world record and estimating c and gamma.

First, we need to clean some data.

## Reading and Cleaning Data

After downloading the data from Wikipedia, I did some cleaning to get it in the 
proper format.

```{r}
#| output: false
library(tidyverse)

raw_data <- read_csv(here::here("Week_1", "hw_1_data.csv")) |>
  select(-9)

clean_df <- function(raw_df) {

  clean_df <- raw_df |>
    mutate(class = case_when(str_detect(Event, "kg") ~ Event, TRUE ~ NA_character_)) |>
    fill(class, .direction = "down") |>
    drop_na() |>
    mutate(Record = str_extract(Record, "[0-9]+") |> as.numeric(),
           class = str_extract(class, "[0-9]+") |> as.numeric())

  clean_df

}

cleaned_data <- raw_data |> clean_df()

```

Let's see what the fits look like:
 
```{r}
#| echo: false
#| warning: false
#| message: false
cleaned_data |>
  ggplot(aes(class, Record, col = Type)) + geom_point() + geom_smooth(method = "lm") + 
  facet_wrap(Event ~ ., scales = "free", ncol =1) + theme_minimal() +
  labs(title = "Linear Fits for Class vs Record", subtitle = "Fits for each Event")
```

Let's do some estimates, one linear model for each sex and event combination:

```{r}
nested_mods <- cleaned_data |>
  mutate(Record = str_extract(Record, "[0-9]+") |> as.numeric(),
         class = str_extract(class, "[0-9]+") |> as.numeric()) |>
  group_nest(Type, Event) |>
  mutate(mod = map(data, ~ lm(log10(Record) ~ log10(class), data = .) |> broom::tidy())) |>
  select(-data) |>
  unnest(mod)

coeffs <- nested_mods |>
  select(Type:estimate) |>
  pivot_wider(names_from = "term", values_from = "estimate") |>
  rename(c = 3, beta = 4) |>
  mutate(c = 10^c) ## we need to exponentiate because the coefficient from the model is on the log scale

coeffs |>
  knitr::kable()

```

```{r}
cleaned_data |>
  left_join(coeffs) |>
  mutate(pred = c*class^beta) |>
  mutate(normalized_record = 100*(Record/((c*class^beta) - 1))) |>
  arrange(desc(normalized_record)) |>
  head() |>
  select(Type, Event, Athlete, normalized_record) |> 
  knitr::kable()
```

